LME-MPM applied to quasi-brittle fracture.

### 2020-05-27 2020_EFM_MPM_Eigensoftening.tex
Cambios figuras y algoritmos
PNA+DM_20_05_27

\newcommand{\DM}[1]{
  \textcolor{gree}{{#1}}
}

@@

to a compromise between two competing objectives, a Pareto set is

@@

\subsection{Fracture modelling approach}
\label{sec:2.3}
Within the context of \acrshort{mpm} formulation, fracture can be
modelled by failing particles according to a suitable failure
criterion. When material points are failed, they are assumed to have
null stress tensor. To reproduce this behaviour in the present
research, the eigensoftening algorithm is introduced in the 
\acrshort{mpm} framework as an alternative approach to the decohesion
model \cite{Zhenmao_2005,Schreyer_2002}. The eigensoftening concept
was originally developed by Navas {\it et al.}
(2017)\cite{Navas_2017_ES} as an extension for quasi-brittle materials
of the eigenerosion proposed by Pandolfi \& Ortiz
(2012)\cite{Pandolfi_2012} for fracture of brittle materials. A
comparison between both in \cite{Navas_2017_ES} shows that the
eigenerosion algorithm significantly overestimates the tensile  
stress and the strain peaks, meanwhile eigensoftening captures the forces and crack
patterns accurately. On the other hand, eigensoftening algorithm agrees
very well with experimental results in all the aspects. Furthermore,
the accuracy of this algorithm has been proven for complex fracture
patterns such as the ones obtained with fiber reinforces concrete (FRC),
\cite{Navas_2018_ES}.\\

@@

attained. For intermediate values a damage value between zero
and one represents the extension to which the material has failed.

@@

theory for brittle fracture was considered by authors
\cite{Navas_2017_ES}. With that purpose, the variation of the averaged

@@

where $m_p$ is the total mass of the $\epsilon$-neighborhood, defined similarly than in the eigenerosion methodology (Eq.~\eqref{eq:mass-EE}).

@@

This definition of the equivalent critical stress leads to a

@@

opening $w_n$ within the band width, $h_{\epsilon}$. Therefore, 

@@

\subsection{$\epsilon$-neighborhood reconstruction : A node-linked method}
\label{sec:epsil-neighb-reconst}
\PNA{The construction of the $\epsilon$-neighborhood is a major issue of the proposed methodology.} Since this operation could be extremely demanding, optimal
numerical implementation should be employed. In opposite with the Cell-linked
method proposed by Allen \& Tildesley \cite{Allen_et_al_1989} where
the definition of a numerical tolerance is required for auxiliar
algorithms like Crossing Number \cite{Shimrat_1962} or Correct
Even-Odd \cite{Galetzka_et_al_2017}, \PNA{the proposed methodology adopted herein is linked to the shape function neighbourhood. In order to avoid the aforementioned error} prone
techniques, we endeavour to exploit the meshfree benefits of
the \acrshort{lme} approximants to introduce a node-linked method. Due
to the global definition of this interpolation technique, the classical
interpretation of the \acrshort{fem} element typically adopted in
\acrshort{mpm} simulations is not required anymore. A linked
list of tributary nodes is reconstructed with the closest node as
starting point at each time step instead. Therefore, each node is assigned to an
unique linked-list of those particles which are closest to it. This
approach allows to reconstruct easily a search list with the support
of the initial connectivity of the mesh and offers a suitable basis
for an easy implementation of other interpolation techniques with
minimum coding effort. A detailed explanation of the proposed
algorithm can be found in \ref{sec:node-linked-method-1}.

@@

As noticed in \cite{Zhang_EE_2020}, the EigenMPM have two important
shortcomings. The first is the presence of stress instabilities even
with \acrshort{gimp} shape functions. To overcome it, \acrshort{lme}
approximants are introduced as an
alternative to the existing interpolation techniques. Its
impressive performance mitigating spurious stress oscillations
\cite{Wobbes2020} under the \acrshort{mpm} framework help to enhance
notoriously the quality of the results as we can see in Section
\ref{sec:3.1}. Where both interpolation techniques
are compared by carry out an eigenerosion simulation. The second
limitation is concerning its inability to simulate properly quasi-brittle
fracture \cite{Navas_2018_ES}. It can be solved through the
eigensoftening algorithm described in Section \ref{sec:2.3}. A proof
of it is exposed in Section \ref{sec:3.3}, where experimental results
of a drop-weight impact test are compared with those produced by
eigensoftening computations.

@@

This problem here presented is devoted to compare how much \acrshort{lme}
approximants are able to improve the result \textit{versus} standard
linear interpolation. It consists in a square plate of size $H = 1$ containing
an initial edge crack of length $0.25 \cdot H$ loaded in a pure mode I
by displacement control on the outer flanks of the plate, Figure
\ref{fig:geometry-cracked-panel-mode-I}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.4\textwidth]{Figure-Mode_I}
  \caption{Geometry and boundary condition of the drop-weight impact test.}
  \label{fig:geometry-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The constitutive model consider for numerical experiment is a
linear-elastic Hookean material, where Young's modulus \gls{E} = 1.06, the
Poisson's ratio \gls{nu} = 0.333, and critical energy-release rate
\gls{Gf} = 0.0001. This set of parameters was taken from \cite{Pandolfi_2012}. In
\acrshort{mpm}, two different discretization are required. In one hand
a cartesian grid of nodes is consider with a nodal spacing value of
0.025. On the other hand, the plate will be modeled with a initial
layout of four particles per element and occupying the Gauss quadrature
positions.\\

Figure \ref{fig:Reactions-cracked-panel-mode-I} clearly shows the
presence of wiggles in both loading and fracture stages
in the reaction-displacements curve when linear interpolation
technique is used. In contrast, \acrshort{lme} simulation does not
exhibit these spurious oscillations and remains linear. Besides that,
linear interpolation produces stiffer results than
\acrshort{lme}. This can be attributed to imprecision in the stress
field due to grid-crossing phenomena.\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{Figure-Reactions-Mode-I}
  \caption{Evolution of the reaction forces plotted \textit{versus}
    the imposed displacement for linear interpolation technique and
    the \acrshort{lme} approximants.}
  \label{fig:Reactions-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
On the other hand, Figure \ref{fig:Stress-cracked-panel-mode-I}
exhibits more clearly the difference between linear interpolation and
\acrshort{lme}, since numerical instability becomes rather significant
when particles crosses the boundary of the element in the linear
case. Although this behavior does not affect significantly to the peak
stress value in a hookean material \cite{Zhang_EE_2020}, if
hyperelastic materials are consider for future simulations,
inaccuracies in the stress field could affect dramatically to the
final result. On the other hand \acrshort{lme} produces soft stress
field evolution even after breaking.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\subfigure[t = 25]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0100-mode-I.png}}
\subfigure[t = 60]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0250-mode-I.png}}
\subfigure[t = 100]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0398-mode-I.png}}
\caption{Evolution of the stress tensor magnitude for linear
  interpolation (pictures in the left side), and \acrshort{lme}
  (pictures in the right side). Both simulations performed with an
  eigenerosion algorithm.}
\label{fig:Stress-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
An important consideration regarding the presence of oscillations once
both parts of the panel are separated, Figure
\ref{fig:Reactions-cracked-panel-mode-I}. These phenomena should not be
attributed to the eigenerosion algorithm since the fracture process is
over. They are due to the dynamic nature of the solver.

\subsection{Drop-weight impact test}
\label{sec:3.2}

Once the accuracy of the \acrshort{lme} approximants have been proven
in Section \ref{sec:3.1}, henceforth will be adopted without comparing
with linear interpolation. Therefore this section is devoted to proof
the accuracy of the eigensoftening algorithm to simulate the behavior
of quasi-brittle materials under dynamic loading. An example of this
load case is the three-point bending test on a notched beam is
conducted under impact loading. This kinds of experiments are
challenging to reproduce numerically with an explicit solver since
fracture occurs in a period of milliseconds and high strength
materials are involved. Accordingly, really small time steps are
required to achieve a simulation numerically stable. Additionally,
this kind of simulations traditionally require the implementation of
contact algorithm to reproduce the hammer impact in the beam. In the
\acrshort{mpm}, the implementation of this feature can be bypassed to
obtain preliminary results. Nevertheless, this kind of algorithm could
be implemented to get ever more accurate solutions. Although this is
out of the scope of this document, future research of authors will overcome this
limitation.\\

As in the study of Zhang {\it et al.} \cite{Zhang_2009,Zhang_2010a},
an impact hammer of 120.6 kg was employed to drop it at an impact
speed of 881 mm/s. The beam dimensions where 100 mm x 100 mm (B x D)
in cross section, and 420 mm in total lenght (L). The initial
notch-depth ratio was approximately 0.5, and the span, S, was fixed at
300 mm during the tests, see Figure
\ref{fig:geometry-drop-weight-impact-test}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{./Figure-impact-test}
  \caption{Geometry and boundary condition of the drop-weight impact test.}
  \label{fig:geometry-drop-weight-impact-test}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The material adopted for the simulation was characterised by Navas
{\it et al.}~\cite{Navas_2017_ES}, and the material properties, such
as the material density, \gls{rho}, the compressive strenght, \gls{fc},
the tensile strenght \gls{ft}, the specific fracture energy, \gls{Gf},
and the elastic modulus \gls{E} are given in table \ref{tab:concrete-properties}.\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{table}
  \centering
  \begin{tabular}[]{c c c c c c c}
    \hline
      &   $\rho$   & $\text{f}_c$ & $\text{f}_t$ & $G_F$ &   E   & $\text{d}_{max}$ \\
      & (kg/m$^3$) &     (MPa)    &     (MPa)    & (N/m) & (GPa) & (mm) \\
    \hline
Value &    2368    &     102.7    &      5.4     &  141  &   31  &  12 \\
    \hline
  \end{tabular}
  \caption[Mechanical properties of thje concrete]{Mechanical
    properties of the high strenght concrete.}
  \label{tab:concrete-properties}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
For the present research, a 2D setting of the experiment has been
adopted. Both the hammer and the concrete beam are explicitly
represented. Several levels of discretization were required to assess
the objectiveness of the obtained results. The results here exposed
are from a discretization of 30616 material points and 6297 nodes. To
achieve better results, an unestructured grid layout has been adopted
focusing the minimum nodal spacing, 0.47 mm, in the middle of the beam and in
the impact surface of the hammer. For this simulation, \gls{gamma-lme} was
fixed to 6.\\

First the reaction and impact forces are validated against their
experimental counterparts. Since the impact forces applied by the
hammer and the reaction forces at the two supports were experimentally
measured, they are compared with numerical ones in Figure
\ref{fig:Reactions-Forces-impact-test}. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{Figure-impact-test-Forces-Time}
  \caption{Evolution of the reaction forces plotted \textit{versus}
    the imposed displacement for linear interpolation technique and
    the \acrshort{lme} approximants.}
  \label{fig:Reactions-Forces-impact-test}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Note that the general trend of both forces are correctly
captured. Observe that in experimental solution the delay between
impact peak and reactions peak does not agree with the numerical
ones. On the other hand, if the delay is computed analytically through
concrete wave speed is much closer to the numerical results. A
feasible explanation to this unexpected result can due the composite
nature of the concrete. Another discrepancy is in the much higher
values of maximum impact and reaction forces in numerical than in
experimental. In this case, this is a consequence of over-rigidity
in the supports. Therefore, this results can be improved in future research
by a suitable contact algorithm.\\

And finally horizontal stress distributions and crack propagation are
extracted, see Figure \ref{fig:Stress-vs-damage-impact-test}. Stress
evolution agrees with the results obtained by \cite{Navas_2017_ES}
under the \acrshort{otm} framework. Two additional considerations about
this results are concerning the interpolation technique. In one hand,
in the opposite to the \acrshort{ugimp} shape function, \acrshort{lme}
allows unstructured grids. This is a remarkable benefit for this kind
of simulations since the majority of the computational effort is
clearly localised in a tiny region of the domain. And in the other
hand there is the possibility to reduce the shape function support to avoid
a material point in one side of the crack is affected by those in the
other side of the crack. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\subfigure[t = 0.27 ms]{\includegraphics[width=0.83\textwidth]{./Figure-impact-test-Damage-vs-Stress-0060}}
\subfigure[t = 0.36 ms]{\includegraphics[width=0.83\textwidth]{./Figure-impact-test-Damage-vs-Stress-0080}}
\subfigure[t = 0.45 ms]{\includegraphics[width=0.83\textwidth]{./Figure-impact-test-Damage-vs-Stress-0100}}
\caption{Evolution of the stress tensor magnitude for linear
  interpolation (pictures in the left side), and \acrshort{lme}
  (pictures in the right side). Both simulations performed with an
  eigenerosion algorithm.}
\label{fig:Stress-vs-damage-impact-test}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Conclusions}
\label{sec:4}
Based on the EigenMPM approach to fracture proposed by Zhang {\it et al.}
\cite{Zhang_EE_2020}, we have proposed two addition improvements which
solves the shortcomings presented in the original concept. First by
bring to the \acrshort{mpm} framework the concept of eigensoftening,
which allow to consider gradual failure process of each material
point. This concept can be compared with a cohesive law in the context
of cohesive fracture. Furthermore, more elaborated damage curves can
be designed to describe the failure process in more sophisticated
materials. In this direction, an interesting future improvement of the
algorithm would be the ability to simulate of crack patterns in
composite material such carbon fiber. This leads a challenging
research topic. Secondly, \acrshort{lme} is employed to remove from
the solution those wiggles due to inaccuracies in the stress
integration. This interpolation technique could help to improve
results in fracture propagation by adapting it through the
deformation gradient. \textit{A priori} this will enhance the already
impressive localisation properties of eigensoftening and eigenerosion
algorithms.

@@

  }  
  \caption{$\epsilon$-neighborhood reconstruction algorithm}
\end{algorithm}

>>>>

\newcommand{\DM}[1]{
  \textcolor{green}{{#1}}
}

@@

\DM{delete to} a compromise between two competing objectives, a Pareto set is

@@

\subsection{Fracture modelling approach}
\label{sec:2.3}
Within the context of \acrshort{mpm} formulation, fracture can be
modelled by failing particles according to a suitable failure
criterion. When material points are failed, they are assumed to have
null stress tensor. To reproduce this behaviour in the present
research, the eigensoftening algorithm is introduced in the 
\acrshort{mpm} framework as an alternative approach to the decohesion
model \cite{Zhenmao_2005,Schreyer_2002}. The eigensoftening concept
was originally developed by Navas {\it et al.}
(2017)\cite{Navas_2017_ES} as an extension for quasi-brittle materials
of the eigenerosion proposed by Pandolfi \& Ortiz
(2012)\cite{Pandolfi_2012} for fracture of brittle materials. \DM{ XXXX I would modify the following  sentence as "The evaluation of both approaches \cite{Navas_2017_ES} show that the
eigenerosion algorithm significantly overestimates the tensile  
stress and the strain peaks, meanwhile eigensoftening captures the forces and crack
patterns accurately XXXXX}. A
comparison between both in \cite{Navas_2017_ES} shows that the
eigenerosion algorithm significantly overestimates the tensile  
stress and the strain peaks, meanwhile eigensoftening captures the forces and crack
patterns accurately. On the other hand, eigensoftening algorithm agrees
very well with experimental results in all the aspects \DM{which aspect?} . Furthermore,
the accuracy of this algorithm has been proven for complex fracture
patterns such as the ones obtained with fiber reinforces concrete (FRC),
\cite{Navas_2018_ES}.\\

@@

attained. \DM{ see the following sentence: comma missing? Clarified--->} For intermediate values a damage value between zero and one represents the extension to which the material has failed. \DM{ hasta acá }

@@

theory for brittle fracture was considered  \DM{ delete by the authors}  by authors
\cite{Navas_2017_ES}. With that purpose \DM{ which purpose?}, the variation of the averaged

@@

where $m_p$ is the total mass of the $\epsilon$-neighborhood, defined similarly  \DM{ Pondría:  defined similarly TO  the eigenerosion method} than in the eigenerosion methodology (Eq.~\eqref{eq:mass-EE}).

@@

This definition \DM{ dos veces definition en la misma oración, podes poner directamente: The equivalent critical stress leads to a definition of the averaged strain energy ... } of the equivalent critical stress leads to a

@@

opening $w_n$ within the band width, $h_{\epsilon}$. \DM{Check band width $h_{\epsilon}$ subindex or superindex? or both. See eq. 18,19 and 20} Therefore, 

@@

\subsection{$\epsilon$-neighborhood reconstruction : A node-linked method}
\label{sec:epsil-neighb-reconst}
\PNA{The construction of the $\epsilon$-neighborhood is a major issue of the proposed methodology.} Since this operation could be extremely demanding, optimal
numerical implementation should be employed. In opposite with the Cell-linked
method proposed by Allen \& Tildesley \cite{Allen_et_al_1989} where
the definition of a numerical tolerance is required for auxiliar
algorithms like Crossing Number \cite{Shimrat_1962} or Correct
Even-Odd \cite{Galetzka_et_al_2017}, \PNA{the proposed methodology adopted herein is linked to the shape function neighbourhood. In order to avoid the aforementioned error} prone
techniques, we endeavour to exploit the meshfree benefits of
the \acrshort{lme} approximants to introduce a node-linked method. Due
to the global definition of this interpolation technique, the classical
interpretation of the \acrshort{fem} element typically adopted in
\acrshort{mpm} simulations is not required anymore. \DM{Intead, a} linked
list of tributary nodes is reconstructed with the closest node as
starting point at each time step \DM{deleted instead}. Therefore, each node is assigned to an
unique linked-list of those particles which are closest to it. This
approach allows to reconstruct easily a search list with the support
of the initial connectivity of the mesh and offers a suitable basis
for an easy implementation of other interpolation techniques with
minimum coding effort. A detailed explanation of the proposed
algorithm can be found in \ref{sec:node-linked-method-1}.

@@

\DM{The proposed approach to overcome the two main shortcomings of EighenMPM \cite{Zhang_EE_2020} has been evaluated with "two benchmark tests"¿?. } \DM{DELETE: As noticed in \cite{Zhang_EE_2020}, the EigenMPM have two important
shortcomings}. The first one deals with the presence of stress instabilities provoked by the spatial discretization, appearing even
when the \acrshort{gimp} shape functions are employed. To overcome it, \acrshort{lme}
approximants are introduced as an
alternative to the existing interpolation techniques. Its
performance in mitigating spurious stress oscillations under the \acrshort{mpm}~\cite{Wobbes2020,Molinos2020} framework helps to enhance
notoriously the quality of the results as it is observed in Section
\ref{sec:3.1}, where both interpolation techniques
are compared by carrying out an eigenerosion simulation. The second
limitation is concerned to its inability to simulate properly quasi-brittle
fracture \cite{Navas_2018_ES}. It can be solved through the
eigensoftening algorithm described in Section \ref{sec:2.3}. A proof
of it is exposed in Section \ref{sec:3.3}, where experimental results
of a drop-weight impact test are compared with experimental ones as well as eigensoftening computations. \PNA{FALTA LO DE BUI}.

@@

The aim of the problem presented hereinafter is to assess the capability of the \acrshort{lme}
approximants to improve the result \textit{versus} the standard linear interpolation. The application consists of a square plate of size $H = 1$ containing
an initial edge crack of length $0.25 \cdot H$ loaded in a pure mode I
by displacement control on the outer flanks of the plate (See Figure
\ref{fig:geometry-cracked-panel-mode-I} for details).
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.4\textwidth]{Figure-Mode_I}
  \caption{Geometry and boundary condition of the drop-weight impact test.}
  \label{fig:geometry-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\PNA{UNIDADES}

\PNA{NOMBRAR ALGO A LOS CASOS ANALITICOS O DE MICHAEL ORTIZ}

The constitutive model considered in the numerical experiment is a
linear-elastic Hookean material, whose Young's modulus \gls{E} = 1.06, the
Poisson's ratio \gls{nu} = 0.333, and critical energy-release rate
\gls{Gf} = 0.0001. This set of parameters has been previously employed by Pandolfi~\textit{et al.}~\cite{Pandolfi_2012}. In
\acrshort{mpm}, two different discretization are required. On the one hand,
a cartesian grid of nodes is considered with a nodal spacing value of
0.025. On the other hand, the plate will be modeled with a initial
layout of four particles per element and occupying the Gauss quadrature
positions.\\

Figure \ref{fig:Reactions-cracked-panel-mode-I} clearly shows the
presence of wiggles in the reaction-displacements curve, in both loading and post-failure stages,
when linear interpolation technique is used. In contrast, \acrshort{lme} simulation does not
exhibit these spurious oscillations and remains linear until the failure. About the strength,
linear interpolation produces stiffer results than
\acrshort{lme}. This can be attributed to imprecision in the stress
field due to grid-crossing phenomena.\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{Figure-Reactions-Mode-I}
  \caption{Evolution of the reaction forces plotted \textit{versus}
    the imposed displacement for linear interpolation technique and
    the \acrshort{lme} approximants.}
  \label{fig:Reactions-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Additionally, Figure \ref{fig:Stress-cracked-panel-mode-I}
exhibits more clearly the difference between linear interpolation and
\acrshort{lme}, since numerical instability becomes rather significant
when particles crosses the boundary of the element in the linear
case, which is observed at 60 seconds. \PNA{Although, apparently, the differences in the obtained peaks are not extremely significant, it can be owing to the aforementioned hookean material employed~\cite{Zhang_EE_2020}. When more sophisticated constitutive models will be employed, severe
%if hyperelastic materials are consider for future simulations,
inaccuracies may appear in the stress field that could affect dramatically to the
final result. Focusing on the post failure behavior, it can be seen how \acrshort{lme} produces soft stress
field evolution even after breaking.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\subfigure[t = 25]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0100-mode-I.png}}
\subfigure[t = 60]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0250-mode-I.png}}
\subfigure[t = 100]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0398-mode-I.png}}
\caption{Evolution of the stress tensor magnitude for linear
  interpolation (pictures in the left side), and \acrshort{lme}
  (pictures in the right side). Both simulations performed with an
  eigenerosion algorithm.}
\label{fig:Stress-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
An important consideration regarding the presence of oscillations once
both parts of the panel are separated is observed in Figure
\ref{fig:Reactions-cracked-panel-mode-I}. These phenomena should not be
attributed to the eigenerosion algorithm since the fracture process is
over. They are due to the dynamic nature of the solver: \PNA{once the energy is release dinamically by the crack, the waves propagate along the domain.}

\subsection{Drop-weight impact test}
\label{sec:3.2}
\PNA{This section is devoted to proof
the accuracy of the eigensoftening algorithm within the \acrshort{mpm} framework when the behavior
of quasi-brittle materials under dynamic loading is simulated. One of the most interesting examples of this
loading case is the three-point bending test on a concrete notched beam when it is conducted under impact loading. Indeed, this experimental test is appropriate to be reproduced numerically with the explicit solver proposed in this paper since
fracture occurs in a period of milliseconds and waves propagate fast along the high strength
materials involved in the aforementioned test.} Accordingly, small time steps are
required to achieve a numerically stable simulation.\\

An interesting case of this experimental test was reported by Zhang {\it et al.} \cite{Zhang_2009,Zhang_2010a}. As in the aforementioned study,
an impact hammer of 120.6 kg has been employed in the proposed simulations to drop it at an impact
speed of 881 mm/s. The beam dimensions were 100 mm x 100 mm (B x D)
in cross section, and 420 mm in total lenght (L). The initial
notch-depth ratio was approximately 0.5, and the span, S, was fixed at
300 mm during the tests (see Figure
\ref{fig:geometry-drop-weight-impact-test}).
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{./Figure-impact-test}
  \caption{Geometry and boundary condition of the drop-weight impact test.}
  \label{fig:geometry-drop-weight-impact-test}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The material adopted for the simulation was characterised by Navas
{\it et al.}~\cite{Navas_2017_ES}, being the material properties, such
as the material density, \gls{rho}, the compressive strenght, \gls{fc},
the tensile strenght \gls{ft}, the specific fracture energy, \gls{Gf},
and the elastic modulus \gls{E}, provided in table \ref{tab:concrete-properties}.\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{table}
  \centering
  \begin{tabular}[]{c c c c c c c}
    \hline
      &   $\rho$   & $\text{f}_c$ & $\text{f}_t$ & $G_F$ &   E   & $\text{d}_{max}$ \\
      & (kg/m$^3$) &     (MPa)    &     (MPa)    & (N/m) & (GPa) & (mm) \\
    \hline
Value &    2368    &     102.7    &      5.4     &  141  &   31  &  12 \\
    \hline
  \end{tabular}
  \caption[Mechanical properties of thje concrete]{Mechanical
    properties of the high strenght concrete.}
  \label{tab:concrete-properties}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

For the present research, a 2D setting of the experiment has been
adopted. Both the hammer and the concrete beam are explicitly
represented in order to properly capture the impact. Several levels of discretization were required to assess
the objectiveness of the obtained results. The results here exposed
are from a discretization of 30616 material points and 6297 nodes. To
achieve better results, an unestructured grid layout has been adopted
focusing the minimum nodal spacing, 0.47 mm, in the middle of the beam and in
the impact surface of the hammer. For this simulation, \gls{gamma-lme} was
fixed to 6. Once the accuracy of the \acrshort{lme} approximants have been proven
in Section \ref{sec:3.1}, this discretization has been adopted.
%without comparing
%with linear interpolation. Therefore \\

First, the reaction and impact forces are validated against their
experimental counterparts. Since the impact forces applied by the
hammer and the reaction forces at the two supports were experimentally
measured, they are compared with numerical ones in Figure
\ref{fig:Reactions-Forces-impact-test}. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{Figure-impact-test-Forces-Time}
  \caption{Evolution of the reaction forces plotted \textit{versus}
    the imposed displacement for linear interpolation technique and
    the \acrshort{lme} approximants.}
  \label{fig:Reactions-Forces-impact-test}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Note that the general trend of both forces are correctly
captured. Observe that in experimental solution the delay between
impact peak and reactions peak does not agree with the numerical
ones. On the other hand, if the delay is computed analytically through
concrete wave speed is much closer to the numerical results. A
feasible explanation to this unexpected result can due the composite
nature of the concrete. Another discrepancy is in the much higher
values of maximum impact and reaction forces in numerical than in
experimental. In this case, this is a consequence of over-rigidity
in the supports. \PNA{Although the traditional \acrshort{mpm} implementation allows collision of two bodies, inaccurate energy conservation is obtained without the implementation of a conservative contact algorithm to reproduce the hammer impact in the beam. %In the
%%\acrshort{mpm}, the implementation of this feature can be bypassed to obtain preliminary results. Nevertheless, this kind of algorithm could
%be implemented to get ever more accurate solutions. 
Despite the observed discrepancies, this is
out of the scope of this document, although future research of authors will overcome this
limitation. Similarly, results of support forces will be more realistic when modelled by a suitable contact algorithm instead of avoiding any vertical displacement}.\\

Finally, longitudinal stress distribution and crack propagation are
depicted in Figure \ref{fig:Stress-vs-damage-impact-test}. Stress
evolution agrees with the results obtained by \cite{Navas_2017_ES}
under the \acrshort{otm} framework. Two additional considerations about
this results are concerning the interpolation technique. On the one hand,
in the opposite to the \acrshort{ugimp} shape function, \acrshort{lme}
allows unstructured grids. This is a remarkable benefit for this kind
of simulations since the majority of the computational effort is
clearly localised in a tiny region of the domain. On the other
hand, there is the possibility to reduce the shape function support to avoid
a material point in one side of the crack is affected by those in the
other side of the crack. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\subfigure[t = 0.27 ms]{\includegraphics[width=0.83\textwidth]{./Figure-impact-test-Damage-vs-Stress-0060}}
\subfigure[t = 0.36 ms]{\includegraphics[width=0.83\textwidth]{./Figure-impact-test-Damage-vs-Stress-0080}}
\subfigure[t = 0.45 ms]{\includegraphics[width=0.83\textwidth]{./Figure-impact-test-Damage-vs-Stress-0100}}
\caption{Evolution of the stress tensor magnitude for linear
  interpolation (pictures in the left side), and \acrshort{lme}
  (pictures in the right side). Both simulations performed with an
  eigenerosion algorithm.}
\label{fig:Stress-vs-damage-impact-test}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Conclusions}
\label{sec:4}
\PNA{Based on the EigenMPM approach to fracture proposed by Zhang {\it et al.}
\cite{Zhang_EE_2020}, two additional improvements have been proposed in the present research in order to 
solve the shortcomings of the original concept that Zhang {\it et al.} pointed out in the aforementioned paper \cite{Zhang_EE_2020}.

Firstly, by
introducing the concept of eigensoftening to the \acrshort{mpm} framework,
which allows to consider gradual failure process of each material point. This concept is analogous the softening law which is traditionally employed in the context
of cohesive fracture in \acrfull{fem}. Furthermore, this failure criterion admits the design of more elaborated damage curves, which can be employed in order to describe the failure process in more sophisticated
materials. In this direction, an interesting future improvement of the
algorithm would be the ability to simulate of crack patterns in
composite material such as steel or carbon fiber reinforced concretes\cite{Ruiz2019}. Interesting researches have been done on this field~\cite{Navas_2018_ES}. In addition, the eigensoftening approach can be extrapolated to rock fracture in the same manner the eigenerosion was~\cite{Wang2017}, taking into account that the volumetric stiffness remains in order to support bulk loads.

Secondly, \acrshort{lme} is employed to remove from
the solution those wiggles provoked by inaccuracies in the stress integration. This spatial discretization was previously employed within the \acrshort{mpm} framework with excellent results under the elastic regime~\cite{Wobbes2020,Molinos2020}. The traditional time scheme of the \acrshort{mpm} has been improved with a Newmark Predictor Corrector scheme as well. In this research, the aforementioned techniques have helped to improve the
results in fracture propagation through the eigensoftening and eigenerosion algorithms. Further improvement may be reached by adapting \acrshort{lme} through the
deformation gradient\MMP{CITA AQUI LOS QUE LO HACEN}. \textit{A priori} this would enhance the localisation properties of the proposed algorithms in more sophisticated domains.

Finally, for better capturing the measurement of forces made in the laboratory, more realistic contact tools have to be developed. Navas~\textif{et al.} demonstrate the importance of the stiffness of the support in order to obtain accurate results of the reaction forces. Interesting contact algorithms within the \acrshort{mpm} methodology are available in the literature \MMP{CITA ALGUNO Y SI SON CONSERVATIVOS EN ENERGIAS, MEJOR}. }

@@

  }
  \label{alg-node-link}
  \caption{$\epsilon$-neighborhood reconstruction algorithm}
\end{algorithm}
